### 2020-02-06 presentations/jsps_2020-03-01/index.Rmd
Update presentation slides
- add inset distribution maps for ecos motifs

```{r load-packages, cache = FALSE}
library(magick)
library(gt)
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(tidyverse)
```

@@

# Load targets
loadd(list = c(
  "taxon_id_map",
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "ja_bioregions_ferns",
  "biodiv_results_ferns",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns"), 
  cache = ja_fern_cache)

@@

```{r richness-plot, out.width = "120%", fig.asp = 0.7}
alpha_div_ferns <-
  alpha_div_ferns %>%
  filter(latitude > 21)

alpha_div_ferns_binned <-
  alpha_div_ferns %>%
  mutate(
    latitude = cut_width(latitude, width = 0.1, closed = "left") %>%
      str_split(",") %>%
      map_chr(1) %>%
      parse_number) %>%
  group_by(latitude) %>%
  summarize_at(
    vars(richness, phy_mpd_obs_z, func_mpd_obs_z), mean, na.rm = TRUE
  )

# Fern richness
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_viridis_c() +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

@@

a + b + plot_layout(widths = c(3,1))

```{r title-slide-image}
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_viridis_c() +
  theme_void() +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    legend.position = "none", 
     panel.background = element_rect(fill = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

@@

```{r mpd-plot, out.width = "120%", fig.asp = 0.7}

# pterido_phy_mpd
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "phy_mpd_obs_z", 
  metric_title = "MPD",
  label = "Mean Phylogenetic Distance") +
  scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
    )) +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

b <- ggplot(
  alpha_div_ferns_binned, 
  aes(y = latitude, x = phy_mpd_obs_z)) +
  geom_path() +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    position = "right") +
  labs(
    subtitle = "",
    x = "MPD"
  ) + 
  annotate(
    "segment", 
    x = 0, xend = 0, 
    y = -Inf, yend = Inf, 
    linetype = 2) + 
  ylim(24, 46) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )

a + b + plot_layout(widths = c(3,1))

@@

```{r ecos-species-prep}
# Make overall dataframe of species contributions to motifs based on theta
species_ranking <-
species_motifs_ferns_6L$theta %>%
  as.data.frame %>%
  rownames_to_column("taxon_id") %>%
  as_tibble %>%
  pivot_longer(-taxon_id, names_to = "motif", values_to = "score") %>%
  left_join(taxon_id_map, by = "taxon_id") %>% 
  select(motif, taxon, score) %>%
  mutate(taxon = str_replace_all(taxon, "_", " ")) %>%
  mutate(score = round(score, 2))

# Define function for making species ranking table
make_species_ranking_table <- function (df, motif_select, slice_n = 15) {

  df %>%
  filter(motif == motif_select) %>%
  arrange(desc(score), taxon) %>% 
  slice(1:slice_n) %>%
  select(taxon, score) %>%
  gt() %>%
  tab_options(
    row.striping.include_table_body = FALSE,
    row.striping.include_stub = FALSE,
    table.background.color = "#000000",
    column_labels.background.color = "#000000") %>%
  tab_style(
    style = list(
      cell_fill(color = "#000000"),
      cell_text(color = "white", size = "medium")
      ),
    locations = cells_column_labels(columns = everything())
    ) %>%
  tab_style(
    style = list(
      cell_fill(color = "#000000"),
      cell_text(color = "white", size = "small")
      ),
    locations = cells_data(
      columns = everything())
  ) %>%
  tab_style(
    style = list(
      cell_text(style = "italic")
      ),
    locations = cells_data(
      columns = vars(taxon))
  )
}

@@

@@

.pull-right-narrow[
```{r ecos-ranking-2}
make_species_ranking_table(species_ranking, 2)
```
]

@@

.pull-right-narrow[
```{r ecos-ranking-3}
make_species_ranking_table(species_ranking, 3)
```
]

@@

.pull-right-narrow[
```{r ecos-ranking-4}
make_species_ranking_table(species_ranking, 4)
```
]

@@

.pull-right-narrow[
```{r ecos-ranking-5}
make_species_ranking_table(species_ranking, 5)
```
]

@@

.pull-right-narrow[
```{r ecos-ranking-6}
make_species_ranking_table(species_ranking, 6)
```
]

@@

.pull-right-narrow[
```{r ecos-ranking-1}
make_species_ranking_table(species_ranking, 1)
```
]

>>>>

```{r load-packages, cache = FALSE}
library(magick)
library(gt)
library(ggtext)
library(drake)
library(knitr)
library(patchwork)
library(scales)
library(scico)
library(magrittr)
library(broom)
library(assertr)
library(sf)
library(tidyverse)
```

@@

# Load targets
loadd(list = c(
  "taxon_id_map",
  "all_cells",
  "alpha_div_ferns",
  "occ_data_ferns",
  "occ_point_data_raw",
  "ja_bioregions_ferns",
  "biodiv_results_ferns",
  "species_motifs_ferns_6L",
  "percent_sex_dip_ferns"), 
  cache = ja_fern_cache)

```{r alpha-div-ferns-prep}
# Add ecostructure motif scores and canape results to alpha diversity,
# filter out a single location south of 21 lat that had no ferns
alpha_div_ferns <-
  species_motifs_ferns_6L$omega %>%
  as.data.frame %>%
  rownames_to_column("site") %>%
  as_tibble %>%
  separate(site, c("longitude", "latitude"), sep = "_") %>%
  left_join(
    mutate_at(alpha_div_ferns, vars(longitude, latitude), as.character),
    by = c("longitude", "latitude")) %>%
  left_join(
    mutate_at(biodiv_results_ferns, vars(longitude, latitude), as.character),
    by = c("longitude", "latitude")) %>%
  mutate_at(vars(longitude, latitude), as.numeric) %>%
  rename_at(vars(1:6 %>% as.character), ~paste0("motif_", .)) %>%
  filter(latitude > 21)
```

```{r title-slide-image, eval = FALSE}
# Code to make title slide image, don't need to run when rendering slides

a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_viridis_c() +
  theme_void() +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    legend.position = "none", 
     panel.background = element_rect(fill = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

@@

```{r richness-plot, out.width = "120%", fig.asp = 0.7}
# Fern richness
# - map
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "richness", 
  metric_title = "No. species ",
  label = "Richness") +
  scale_fill_viridis_c() +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

# - line tracking mean richness
alpha_div_ferns_binned <-
  alpha_div_ferns %>%
  mutate(
    latitude = cut_width(latitude, width = 0.1, closed = "left") %>%
      str_split(",") %>%
      map_chr(1) %>%
      parse_number) %>%
  group_by(latitude) %>%
  summarize_at(
    vars(richness, phy_mpd_obs_z, func_mpd_obs_z), mean, na.rm = TRUE
  )

@@

a + b + plot_layout(widths = c(3,1))
```

@@

```{r mpd-plot, out.width = "120%", fig.asp = 0.7}
# Phylogenetic diversity (MPD)
# - map
a <- make_diversity_map(
  div_data = alpha_div_ferns, 
  occ_data = occ_data_ferns, 
  div_metric = "phy_mpd_obs_z", 
  metric_title = "MPD",
  label = "Mean Phylogenetic Distance") +
  scale_fill_scico(
    palette = "vik", 
    na.value="dark grey",
    limits = c(
      -get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3), 
      get_limit(alpha_div_ferns, phy_mpd_obs_z, "abs", 3)
    )) +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    limits = c(24, 46)) +
  theme(
    plot.subtitle = element_blank())

# line tracking mean MPD
b <- ggplot(
  alpha_div_ferns_binned, 
  aes(y = latitude, x = phy_mpd_obs_z)) +
  geom_path() +
  scale_y_continuous(
    labels = scales::label_number(suffix = "\u00b0N", accuracy = 1),
    position = "right") +
  labs(
    subtitle = "",
    x = "MPD"
  ) + 
  annotate(
    "segment", 
    x = 0, xend = 0, 
    y = -Inf, yend = Inf, 
    linetype = 2) + 
  ylim(24, 46) +
  theme(
    panel.background = element_blank(),
    panel.grid.major = element_line(size = 0.1, color = "dark grey"),
    panel.grid.minor = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y = element_blank()
  )

a + b + plot_layout(widths = c(3,1))
```

@@

```{r ecos-species-prep}
# Make overall dataframe of species contributions to motifs based on theta
species_ranking <-
species_motifs_ferns_6L$theta %>%
  as.data.frame %>%
  rownames_to_column("taxon_id") %>%
  as_tibble %>%
  pivot_longer(-taxon_id, names_to = "motif", values_to = "score") %>%
  left_join(taxon_id_map, by = "taxon_id") %>% 
  select(motif, taxon, score) %>%
  mutate(taxon = str_replace_all(taxon, "_", " ")) %>%
  mutate(score = round(score, 2)) %>%
  group_by(taxon) %>%
  mutate(score_sd = sd(score)) %>%
  ungroup

# Define function for making species ranking table
make_species_ranking_table <- function (ranking_data, motif_select, taxon_select = NULL, slice_n = 10) {

  ranking_data_sliced <- ranking_data %>%
    filter(motif == motif_select) %>%
    # Select top-scoring species with high SD
    # (high SD should indicate mostly contributing to
    # a single motif)
    arrange(desc(score), desc(score_sd)) %>%
    slice(1:slice_n) %>%
    select(taxon, score) %>%
    mutate(score = as.factor(score)) %>%
    # Arrange alphabetically within score
    arrange(desc(score), taxon)

  table <-
    ranking_data_sliced %>%
    gt() %>%
    tab_options(
      row.striping.include_table_body = FALSE,
      row.striping.include_stub = FALSE,
      table.background.color = "#000000",
      column_labels.background.color = "#000000") %>%
    tab_style(
      style = list(
        cell_fill(color = "#000000"),
        cell_text(color = "white", size = "medium")
      ),
      locations = cells_column_labels(columns = everything())
    ) %>%
    tab_style(
      style = list(
        cell_fill(color = "#000000"),
        cell_text(color = "white", size = "small")
      ),
      locations = cells_data(
        columns = everything())
    ) %>%
    tab_style(
      style = list(
        cell_text(style = "italic")
      ),
      locations = cells_data(
        columns = vars(taxon))
    )

  if(!is.null(taxon_select))
    table <-
      table %>%
      tab_style(
        style = list(
          cell_fill(
            color = "red",
            alpha = 1.0
          )
        ),
        locations = cells_data(
          columns = everything(),
          rows = taxon == ranking_data_sliced$taxon[[taxon_select]])
      )

  table

}

# Define function to make simple distribution map for a species
plot_distribution <- function (ranking_data, motif_select, taxon_select, occ_data, show_title = TRUE) {

   ranking_data <- ranking_data %>%
    filter(motif == motif_select) %>%
    arrange(desc(score), desc(score_sd)) %>%
    select(taxon, score) %>%
    mutate(score = as.factor(score)) %>%
    arrange(desc(score), taxon)

  cells_present <-
    occ_data %>%
    filter(taxon_name == ranking_data$taxon[[taxon_select]]) %>%
    select(secondary_grid_code) %>%
    unique %>%
    mutate(present = TRUE)

  plot_data <-
    occ_data %>%
    select(secondary_grid_code, latitude, longitude) %>%
    left_join(cells_present) %>%
    mutate(present = replace_na(present, FALSE))

  plot <- ggplot(plot_data, aes(x = longitude, y = latitude, fill = present)) +
    geom_tile() +
    coord_quickmap(
      xlim = c(pull(occ_data, longitude) %>% min %>% floor,
               pull(occ_data, longitude) %>% max %>% ceiling),
      ylim = c(pull(occ_data, latitude) %>% min %>% floor,
               pull(occ_data, latitude) %>% max %>% ceiling)
    ) +
    scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "grey80"), na.value = "grey80", guide = FALSE) +
    theme_void()

  if(isTRUE(show_title))
    plot <- plot +
    labs(
      title = ranking_data$taxon[[taxon_select]]
    )

  plot
}
```

```{r write_out_motif_maps, eval = FALSE}
# Don't run when making slides, but it's a useful practice to check the
# distributions for the top 10 taxa for each motif
plot_dist_by_motif <- function(species_ranking, motif, occ_data_ferns) {
  dist_maps <- purrr::map(1:10, ~plot_distribution(species_ranking, motif, ., occ_data_ferns)) %>%
    patchwork::wrap_plots(ncol = 2, nrow = 5)

  ggsave(plot = dist_maps, file = glue::glue("motif_{motif}_maps.png"), width = 6, height = 12, units = "in")
}

walk(1:6, ~plot_dist_by_motif(species_ranking, ., occ_data_ferns))
```

@@

.pull-right-narrow[
```{r ecos-motif-1}
make_species_ranking_table(species_ranking, 1)
```
]

.pull-right-narrow[
```{r ecos-motif-1-focus-3}
make_species_ranking_table(species_ranking, 1, 3)
plot_distribution(species_ranking, 1, 3, occ_data_ferns, show_title = FALSE)
```
]

.pull-right-narrow[
```{r ecos-motif-1-focus-7}
make_species_ranking_table(species_ranking, 1, 7)
plot_distribution(species_ranking, 1, 7, occ_data_ferns, show_title = FALSE)
```
]

@@

.pull-right-narrow[
```{r ecos-motif-2}
make_species_ranking_table(species_ranking, 2)
```
]

.pull-right-narrow[
```{r ecos-motif-2-focus-1}
make_species_ranking_table(species_ranking, 2, 1)
plot_distribution(species_ranking, 2, 1, occ_data_ferns, show_title = FALSE)
```
]

.pull-right-narrow[
```{r ecos-motif-2-focus-2}
make_species_ranking_table(species_ranking, 2, 2)
plot_distribution(species_ranking, 2, 2, occ_data_ferns, show_title = FALSE)
```
]

@@

.pull-right-narrow[
```{r ecos-motif-3}
make_species_ranking_table(species_ranking, 3)
```
]

.pull-right-narrow[
```{r ecos-motif-3-focus-3}
make_species_ranking_table(species_ranking, 3, 3)
plot_distribution(species_ranking, 3, 3, occ_data_ferns, show_title = FALSE)
```
]

.pull-right-narrow[
```{r ecos-motif-3-focus-7}
make_species_ranking_table(species_ranking, 3, 7)
plot_distribution(species_ranking, 3, 7, occ_data_ferns, show_title = FALSE)
```
]

@@

.pull-right-narrow[
```{r ecos-motif-4}
make_species_ranking_table(species_ranking, 4)
```
]

.pull-right-narrow[
```{r ecos-motif-4-focus-2}
make_species_ranking_table(species_ranking, 4, 2)
plot_distribution(species_ranking, 4, 2, occ_data_ferns, show_title = FALSE)
```
]

.pull-right-narrow[
```{r ecos-motif-4-focus-8}
make_species_ranking_table(species_ranking, 4, 8)
plot_distribution(species_ranking, 4, 8, occ_data_ferns, show_title = FALSE)
```
]

@@

.pull-right-narrow[
```{r ecos-motif-5}
make_species_ranking_table(species_ranking, 5)
```
]

.pull-right-narrow[
```{r ecos-motif-5-focus-3}
make_species_ranking_table(species_ranking, 5, 3)
plot_distribution(species_ranking, 5, 3, occ_data_ferns, show_title = FALSE)
```
]

.pull-right-narrow[
```{r ecos-motif-5-focus-7}
make_species_ranking_table(species_ranking, 5, 7)
plot_distribution(species_ranking, 5, 7, occ_data_ferns, show_title = FALSE)
```
]

@@

.pull-right-narrow[
```{r ecos-motif-6}
make_species_ranking_table(species_ranking, 6)
```
]

@@

.pull-right-narrow[
```{r ecos-motif-6-focus-1}
make_species_ranking_table(species_ranking, 6, 1)
plot_distribution(species_ranking, 6, 1, occ_data_ferns, show_title = FALSE)
```
]

.pull-right-narrow[
```{r ecos-motif-6-focus-2}
make_species_ranking_table(species_ranking, 6, 2)
plot_distribution(species_ranking, 6, 2, occ_data_ferns, show_title = FALSE)
```
]

## Motif 2 tends to cluster, motif 6 tends to overdisperse

.pull-left[<img src="images/species_motifs_ferns_all_crop.png" alt="drawing">]

.pull-right[
```{r mpd-vs-motifs}
# Plot omega score vs motif
a <- make_scatter_with_lm(alpha_div_ferns, "motif_1", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[1]]) +
  labs(x = "", title = "Motif 1")
b <- make_scatter_with_lm(alpha_div_ferns, "motif_2", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[2]]) +
  labs(title = "Motif 2", y = "", x = "")
c <- make_scatter_with_lm(alpha_div_ferns, "motif_3", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[3]]) +
  labs(title = "Motif 3", y = "", x = "")
d <- make_scatter_with_lm(alpha_div_ferns, "motif_4", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[4]]) +
  labs(title = "Motif 4")
e <- make_scatter_with_lm(alpha_div_ferns, "motif_5", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[5]]) +
  labs(title = "Motif 5", y = "")
f <- make_scatter_with_lm(alpha_div_ferns, "motif_6", "phy_mpd_obs_z", "Score", "MPD", ecos_cols[[6]]) +
  labs(title = "Motif 6", y = "")
a + b + c + d + e + f + plot_layout(ncol = 2, nrow = 3)
```
]
